### 2020-03-28 thesis/content/chapter-ltd.tex
started with k-gnn limits section

As we saw in \cref{sec:ltag:formulation:gcnn:conv}, the constituents of neighborhood convolutions are spanned by \ac{bfs} trees of depth $T$, where $T$ is the number of convolutional layers.
From this perspective the problem of learning constituents corresponds to learning a pruning operator on the branches of \ac{bfs} trees.
Such a pruning operator filters the edges that are traversed in each \ac{bfs} step.
There are two general edge filtering strategies:
\begin{enumerate}[label={\textbf{\arabic*.}}]
	\item \textbf{Edge prefiltering:}
		Here the edge filtering and convolution operations are performed in independent steps:
		First the edges of a given graph are filtered, then the convolution layers are applied to the filtered graph.
		This so-called \textit{edge prefiltering} strategy is illustrated in \cref{fig:ltd:subtree-filtering}.
		The main advantage of prefiltering is that it allows arbitrary combinations of edge filtering and convolution approaches.
		The main disadvantage is however that the same edges are removed in all \ac{bfs} subtrees.
		This restricts the expressive power of the learned decomposition function as shown in \cref{fig:ltd:dynamic-subtree-filtering}.
	\item \textbf{Dynamic edge filtering:}
		By filtering edges as part of the convolution operation itself, more flexible decompositions can be obtained.
		In the \textit{dynamic edge filtering} strategy the edge filter is part of the convolution operation and decides which neighbors of a given root node should be aggregated.
		Using this strategy a decomposition such as that shown in \cref{fig:ltd:dynamic-subtree-filtering} is learnable.
\end{enumerate}
\begin{figure}[ht]
	\centering
	\includegraphics[width=0.8\linewidth]{gfx/ltd/subtree-filtering.pdf}
	\caption{Illustration of a pruned \ac{bfs} subtree when two edges are removed via prefiltering.}\label{fig:ltd:subtree-filtering}
\end{figure}
\begin{figure}[ht]
	\centering
	\includegraphics[width=0.75\linewidth]{gfx/ltd/dynamic-subtree-filtering.pdf}
	\caption[A graph decomposition obtained via dynamic edge filtering which cannot be modeled by edge prefiltering.]{
		A graph decomposition obtained via dynamic edge filtering which cannot be modeled by edge prefiltering.
		Depending on the \ac{bfs} root node $v_{\mathrm{root}}$ different edges are removed.
		For $v_{\mathrm{root}} \in \textcolor{t_blue}{\{ v_3, v_4 \}}$ the edges $\{ (v_3, v_1), (v_3, v_2), (v_4, v_5), (v_4, v_6) \}$ are filtered out while for $v_{\mathrm{root}} \in \textcolor{t_red}{\{ v_1, v_2, v_5, v_6 \}}$ the edge $(v_3, v_4)$ is removed. % chktex 25
	}\label{fig:ltd:dynamic-subtree-filtering}
\end{figure}
Irrespective of the chosen edge filtering strategy, the decision whether to remove a given edge $e_{ij} = (v_i, v_j)$ or not has to be made based on relevant information about $e_{ij}$ and its surroundings.
One natural way to encode the information about $e_{ij}$ is via an edge feature vector $z_{ij} \in \mathbb{R}^d$ that is fed into the edge filter.
We propose that those edge feature vectors could be obtained by a 2-\acs{wl} inspired graph convolution layer since 2-\acs{wl} colors represent the structural roles of edges as described in \cref{sec:related:character:wl}.

\section{Shortcomings of the Existing $k$-\acs*{gnn}}%
\label{sec:ltd:kgnn-problems}

>>>>

As we saw in \cref{sec:ltag:formulation:gcnn:conv}, the constituents of neighborhood convolutions are spanned by \ac{bfs} trees of depth $T$, where $T$ is the number of convolutional layers.
From this perspective the problem of learning constituents corresponds to learning a pruning operator on the branches of \ac{bfs} trees.
Such a pruning operator filters the edges that are traversed in each \ac{bfs} step.
There are two general edge filtering strategies:
\begin{enumerate}[label={\textbf{\arabic*.}}]
	\item \textbf{Edge prefiltering:}
		Here the edge filtering and convolution operations are performed in independent steps:
		First the edges of a given graph are filtered, then the convolution layers are applied to the filtered graph.
		This so-called \textit{edge prefiltering} strategy is illustrated in \cref{fig:ltd:subtree-filtering}.
		The main advantage of prefiltering is that it allows arbitrary combinations of edge filtering and convolution approaches.
		The main disadvantage is however that the same edges are removed in all \ac{bfs} subtrees.
		This restricts the expressive power of the learned decomposition function as shown in \cref{fig:ltd:dynamic-subtree-filtering}.
	\item \textbf{Dynamic edge filtering:}
		By filtering edges as part of the convolution operation itself, more flexible decompositions can be obtained.
		In the \textit{dynamic edge filtering} strategy the edge filter is part of the convolution operation and decides which neighbors of a given root node should be aggregated.
		Using this strategy a decomposition such as that shown in \cref{fig:ltd:dynamic-subtree-filtering} is learnable.
\end{enumerate}
\begin{figure}[ht]
	\centering
	\includegraphics[width=0.8\linewidth]{gfx/ltd/subtree-filtering.pdf}
	\caption{Illustration of a pruned \ac{bfs} subtree when two edges are removed via prefiltering.}\label{fig:ltd:subtree-filtering}
\end{figure}
\begin{figure}[ht]
	\centering
	\includegraphics[width=0.75\linewidth]{gfx/ltd/dynamic-subtree-filtering.pdf}
	\caption[A graph decomposition obtained via dynamic edge filtering which cannot be modeled by edge prefiltering.]{
		A graph decomposition obtained via dynamic edge filtering which cannot be modeled by edge prefiltering.
		Depending on the \ac{bfs} root node $v_{\mathrm{root}}$ different edges are removed.
		For $v_{\mathrm{root}} \in \textcolor{t_blue}{\{ v_3, v_4 \}}$ the edges $\{ (v_3, v_1), (v_3, v_2), (v_4, v_5), (v_4, v_6) \}$ are filtered out while for $v_{\mathrm{root}} \in \textcolor{t_red}{\{ v_1, v_2, v_5, v_6 \}}$ the edge $(v_3, v_4)$ is removed. % chktex 25
	}\label{fig:ltd:dynamic-subtree-filtering}
\end{figure}
Irrespective of the chosen edge filtering strategy, the decision whether to remove a given edge $e_{ij} = (v_i, v_j)$ or not has to be made based on relevant information about $e_{ij}$ and its surroundings.
One way to encode the information about $e_{ij}$ is via an edge feature vector $z_{ij} \in \mathbb{R}^d$ that is fed into the edge filter.

We propose that those edge feature vectors can be obtained via a 2-\acs{wl} inspired graph convolution layer since 2-\acs{wl} colors naturally represent the structural roles of edges as described in \cref{sec:related:character:wl}.
We already saw a \ac{gcnn} architecture in \cref{sec:related:gcr:nn} which could potentially compute such 2-\acs{wl} inspired edge feature vectors, the $k$-\acs{gnn}~\cite{Morris2019}.
Its variant for $k = 2$ produces edge feature vectors that can be used as the input for an edge filter.
However, as we will see in the next section, there are significant limitations to the discriminative and computational power of 2-\acsp{gnn}.

\section{Limitations of the Existing 2-\acs*{gnn}}%
\label{sec:ltd:kgnn-limits}

The $k$-\acs{gnn} is a \ac{gcnn} inspired by the $k$-\ac{wl} algorithm, it convolves feature vectors of vertex $k$-multisets.
In this section we will compare its $k = 2$ variant with the 2-\ac{wl} algorithm.
The main difference between the two boils down to their notion of ``neighborhood''.
As already briefly mentioned in \cref{sec:related:gcr:nn}, 2-\acsp{gnn} define the neighbors of an edge $\textcolor{t_blue}{e_{ij}} = (\textcolor{t_red}{v_i}, \textcolor{t_darkgreen}{v_j})$ to be the edges that are incident to either $\textcolor{t_red}{v_i}$ or $\textcolor{t_darkgreen}{v_j}$. % chktex 25
In 2-\acs{wl} on the other hand, the neighbors of $\textcolor{t_blue}{e_{ij}}$ are the edge pairs ${\left\{(\textcolor{t_red}{e_{il}}, \textcolor{t_darkgreen}{e_{lj}})\right\}}_{v_l \in \mathcal{V}_G}$, i.e.\ all possible paths of length two that start at $\textcolor{t_red}{v_i}$ and end at $\textcolor{t_darkgreen}{v_j}$. % chktex 21 chktex 25
This difference becomes clear when comparing the definition of convolution in 2-\acsp{gnn} with that of color refinement in 2-\acs{wl} (see \cfullref[ on ]{eq:related:kgnn-layer} and \cfullref[ on ]{defn:related:wlk-refine}):
\begin{alignat*}{4}
	\text{2-\acs{gnn}: } && Z^{(t)}[e_{ij}] &= \sigma\Biggl( \textcolor{t_blue}{Z^{(t-1)}[e_{ij}]} W^{(t)} + &&\left( \smashoperator[r]{\sum_{v_l \in \Gamma_G(v_j)}} \textcolor{t_red}{Z^{(t-1)}[e_{il}]} + \smashoperator[lr]{\sum_{v_l \in \Gamma_G(v_i)}} \textcolor{t_darkgreen}{Z^{(t-1)}[e_{lj}]} \right) W_{\Gamma}^{(t)} \Biggr) \\ % chktex 25
	\text{2-\acs{wl}: } && \chi_{G,2}^{(t)}(e_{ij}) &= h\Bigl(\textcolor{t_blue}{\chi_{G,2}^{(t-1)}(e_{ij})}, &&\ldblbrace (\textcolor{t_red}{\chi_{G,2}^{(t-1)}(e_{il})}, \textcolor{t_darkgreen}{\chi_{G,2}^{(t-1)}(e_{lj})})\, |\, v_l \in \mathcal{V}_G \rdblbrace\Bigr) % chktex 25
\end{alignat*}
We will now analyze what those different notions of neighborhood imply for the discriminative and computational power of 2-\acsp{gnn}.
In the first step we show that the discriminative power of 2-\acsp{gnn} on all graphs $G \in \mathcal{G}$ is upper bounded by that of 1-\acs{wl} on the so-called \textit{edge neighborhood graphs} $G_{\mathcal{E}} \in \mathcal{G}_{\mathcal{E}}$.
\begin{defn}
	The \textit{edge neighborhood graph} of a given graph $G = (\mathcal{V}_G, \mathcal{E}_G)$ is defined as $G_{\mathcal{E}} \coloneqq (\mathcal{V}_G \cup \mathcal{E}_G, \mathcal{E}_{G_{\mathcal{E}}})$. Its \textit{edge neighborhood edges} are defined as
	\begin{align*}
		\mathcal{E}_{G_{\mathcal{E}}} \coloneqq
		\{ s \in \{ (v, e), (e, v) \}\, |\, v \in \mathcal{V}_G \land e \in \Gamma_G(v) \} \cup \{ (e, e') \in \mathcal{E}_G^2\, |\, e \cap e' \neq \emptyset \} % chktex 21
		\text{.}
	\end{align*}
\end{defn}
\begin{prop}
	For all 2-\acsp{gnn} $h_2$ there is a \ac{gin} $h_1$ s.t.\ $\forall G \in \mathcal{G}: h_2(G) = h_1(G_{\mathcal{E}})$.
\end{prop}
\begin{proof}
	Note that the 2-\acs{gnn} convolution operation defined in \cfullref[ on ]{eq:related:kgnn-layer} can be rewritten as
	\begin{align*}
		Z^{(t)}[e] = \sigma \left( Z^{(t-1)}[e] W^{(t)} + \smashoperator[lr]{\sum_{e' \in \Gamma_{G_{\mathcal{E}}}(e)}} Z^{(t-1)}[e'] W_{\Gamma}^{(t)} \right)
	\end{align*}
\end{proof}

### 2020-03-28 thesis/content/chapter-related-work.tex
started with k-gnn limits section

\subsubsection{Spatial \acp{gcnn}}
Looking at the \ac{gcn} layer defined in \cref{eq:related:gcn-layer}, notice that, even though it is motivated by spectral graph theory, the spectrum $\lambda_G$ is not actually directly used.
This allows for a very different perspective on \acp{gcn}:
Instead of interpreting convolutions as applications of linear spectral filters, we can interpret them as vertex neighborhood aggregations, similar to 1-\acs{wl}.
From this perspective a \ac{gcn} can be seen as a so-called spatial \ac{gcnn} because it operates directly in the vertex domain.
To see why this is the case, we rewrite \cref{eq:related:gcn-layer} and compare it with the 1-\acs{wl} color refinement strategy (see \cfullref{defn:related:wl1-refine}):
\begin{alignat*}{2}
	\text{\ac*{gcn}:}\quad&& Z^{(t)}[v] &= \sigma\left(\left( {\mu(v)}^2 Z^{(t-1)}[v] + \smashoperator[lr]{\sum_{u \in \Gamma_G(v)}} \mu(v) \mu(u) Z^{(t-1)}[u] \right) \Theta^{(t)} \right) \\
	\text{1-\acs{wl}:}\quad&& \chi_{G,1}^{(t)}(v) &= h\left( \chi_{G,1}^{(t-1)}(v), \ldblbrace \chi_{G,1}^{(t-1)}(u)\, |\, u \in \Gamma_{G}(v) \rdblbrace\right)
\end{alignat*}
Here $\mu(v) \coloneqq {\left(\left| \Gamma_G(v) \right| + 1\right)}^{-\frac{1}{2}}$ are normalization factors introduced by $\tilde{D}_G^{-\frac{1}{2}}$.
The comparison shows that \ac{gcn} convolutions can be understood as iterative refinements of continuous ``color vectors'' $Z^{(t)}[v] \in \mathbb{R}^{d^{(t)}}$.
Then the continuous analogue to the injective 1-\acs{wl} hash function $h: \mathcal{C}^{*} \to \mathcal{C}$ can be defined as $h_{\mathrm{\acs*{gcn}}}(z_0, z_1, \dots, z_m) = \sigma(\sum_{j=0}^m \mu(v_0) \mu(v_j) z_j \Theta^{(t)})$ which computes a single color vector from multiple color vectors $z_j$ where $v_j$ are the corresponding vertices.

@@

\paragraph{Higher dimensional \ac{wl} \acp{gnn}}
The majority of \ac{gnn} methods uses a neighborhood aggregation scheme which limits their discriminative and, more importantly, computational power to that of 1-\acs{wl} (see \cfullref{defn:related:wl-count-detect}).
To go beyond the limits of 1-\acs{wl}, \citet{Morris2019} proposed the first \ac{gcnn} architecture inspired by the higher dimensional $k$-\acs{wl} algorithms.
Similar to the $k$-LWL kernel which we looked at in \cref{sec:related:gcr:kernel}, their so-called $k$-\acs{gnn} uses the $k$-multiset and neighborhood localization optimizations to keep runtime costs feasible.
The $k$-\ac{gnn} convolution is described by
\begin{align}
	Z^{(t)}[s] \coloneqq \sigma\left( {Z^{(t-1)}[s]} W^{(t)} + \smashoperator[lr]{\sum_{v \in s, u \in \Gamma_G(v)}} {Z^{(t-1)}[s \setminus \{ v \} \cup \{ u \}]} W_{\Gamma}^{(t)} \right) \text{.} \label{eq:related:kgnn-layer}
\end{align}
It convolves the feature vectors of $k$-multisets $s \subseteq \mathcal{V}_G$ instead of single vertices $v \in \mathcal{V}_G$.
Apart from this fundamental difference to 1-\acs{wl} bounded \acp{gnn}, $k$-\acp{gnn} additionally use two learnable parameter matrices $W^{(t)}, W_{\Gamma}^{(t)} \in \mathbb{R}^{d^{(t-1)} \times d^{(t)}}$ instead of the single $\Theta^{(t)}$ used in \acp{gcn}.
This separation of the parameter matrix effectively fulfills the same purpose as the \ac{mlp} in \ac{gin}; it makes the implicitly defined $k$-\acs{gnn} hash function injective.
This can be seen by realizing that a stack of $k$-\acs{gnn} convolution layers with $W_{\Gamma}^{(t)} = 0$ for all but the first layer simulates a \ac{mlp}.

As a final remark, note that the concept of ``neighborhood'' in \cref{eq:related:kgnn-layer} is slightly different from that used by the $k$-\acs{wl} algorithm (see \cfullref{defn:related:wlk-refine}):
$k$-\acs{gnn} considers a single $k$-multiset as a neighbor while $k$-\acs{wl} uses $k$-sets of $k$-multisets as neighbors.
We will get back to this difference later.

@@

The graph convolution approaches that we just looked at all produce a set of convolved feature vectors ${\left\{ Z^{(T)}[s_i] \right\}}_{i=1}^{m}$ after $T$ convolutional layers. % chktex 21
For 1-\acs{wl} bounded architectures like \ac{gcn} or \ac{gin}, there is one vector for each of the $m = {|\mathcal{V}_G|}$ vertices, while a $k$-\acs{wl} inspired architecture like $k$-\acs{gnn} produces $m \leq {|\mathcal{V}_G|}^{k}$ output vectors.
In order to solve the \ac{gcr} problem, those vector sets need to be combined into a final predicted class or regression value.
To do this, so-called \textit{graph pooling layers} are used.
Generally speaking there are two types of pooling approaches:
\begin{enumerate}[label=\textbf{\arabic*.}]
	\item \textbf{Hierarchical Pooling:}
		This type of pooling is similar to that found in \acp{cnn} on pictures where the resolution is iteratively reduced.
		Similarly hierarchical graph pooling iteratively \textit{coarsens} a graph every couple of convolutional layers.
		Graph coarsening works by merging vertices in a spectrum-preserving manner~\cite{Loukas2018}.
		This is done until a single merged vertex is left whose feature vector then represents the entire graph.
	\item \textbf{Global Pooling}:
		Alternatively pooling can also be performed in a single step after the convolutional layers.
		This type of pooling takes all $m$ feature vectors and directly maps them to a single graph feature vector.
\end{enumerate}
In this thesis we will focus on global pooling.
The most simple global pooling approaches use static aggregation functions like component-wise $\min$, $\max$ or $\mean$ to produce a single graph feature vector $z_G \in \mathbb{R}^{d^{(T)}}$.
This vector is then fed into a standard \ac{mlp} to produce the final prediction.
The state-of-the-art graph pooling approaches go beyond static aggregation and try to incorporate structural information.
We will now briefly look at two such approaches.

>>>>

\subsubsection{Spatial \acp{gcnn}}
Looking at the \ac{gcn} layer defined in \cref{eq:related:gcn-layer}, notice that, even though it is motivated by spectral graph theory, the spectrum $\lambda_G$ is not actually directly used.
This allows for a very different perspective on \acp{gcn}:
Instead of interpreting convolutions as applications of linear spectral filters, we can interpret them as vertex neighborhood aggregations, similar to 1-\acs{wl}.
From this perspective a \ac{gcn} can be seen as a so-called spatial \ac{gcnn} because it operates directly in the vertex domain.
To see why this is the case, we rewrite \cref{eq:related:gcn-layer} and compare it with the 1-\acs{wl} color refinement strategy (see \cfullref{defn:related:wl1-refine}):
\begin{alignat*}{2}
	\text{\ac*{gcn}:}\quad&& Z^{(t)}[v] &= \sigma\left(\left( {\mu(v)}^2 \textcolor{t_blue}{Z^{(t-1)}[v]} + \smashoperator[lr]{\sum_{u \in \Gamma_G(v)}} \mu(v) \mu(u) \textcolor{t_red}{Z^{(t-1)}[u]} \right) \Theta^{(t)} \right) \\ % chktex 25
	\text{1-\acs{wl}:}\quad&& \chi_{G,1}^{(t)}(v) &= h\left( \mkern60mu\textcolor{t_blue}{\chi_{G,1}^{(t-1)}(v)}, \ldblbrace \textcolor{t_red}{\chi_{G,1}^{(t-1)}(u)}\, |\, u \in \Gamma_{G}(v) \rdblbrace\right) % chktex 25
\end{alignat*}
Here $\mu(v) \coloneqq {\left(\left| \Gamma_G(v) \right| + 1\right)}^{-\frac{1}{2}}$ are normalization factors introduced by $\tilde{D}_G^{-\frac{1}{2}}$.
The comparison shows that \ac{gcn} convolutions can be understood as iterative refinements of continuous ``color vectors'' $Z^{(t)}[v] \in \mathbb{R}^{d^{(t)}}$.
Then the continuous analogue to the injective 1-\acs{wl} hash function $h: \mathcal{C}^{*} \to \mathcal{C}$ can be defined as $h_{\mathrm{\acs*{gcn}}}(z_0, z_1, \dots, z_m) = \sigma(\sum_{j=0}^m \mu(v_0) \mu(v_j) z_j \Theta^{(t)})$ which computes a single color vector from multiple color vectors $z_j$ where $v_j$ are the corresponding vertices.

@@

\paragraph{Higher dimensional \ac{wl} \acp{gnn}}
The majority of \ac{gnn} methods uses a neighborhood aggregation scheme which limits their discriminative and, more importantly, computational power to that of 1-\acs{wl} (see \cfullref{defn:related:wl-count-detect}).
To go beyond the limits of 1-\acs{wl}, \citet{Morris2019} proposed the first \ac{gcnn} architecture inspired by the higher dimensional $k$-\acs{wl} algorithms.
Similar to the $k$-LWL kernel which we looked at in \cref{sec:related:gcr:kernel}, their so-called $k$-\acs{gnn} uses the $k$-multiset and neighborhood localization optimizations to keep runtime costs feasible.
The $k$-\ac{gnn} convolution is described by
\begin{align}
	Z^{(t)}[s] \coloneqq \sigma\left( {Z^{(t-1)}[s]} W^{(t)} + \smashoperator[lr]{\sum_{{v \in s}, {u \in \Gamma_G(v)}}} {Z^{(t-1)}[s \setminus \{ v \} \cup \{ u \}]} W_{\Gamma}^{(t)} \right) \text{.} \label{eq:related:kgnn-layer}
\end{align}
It convolves the feature vectors of $k$-multisets $s \subseteq \mathcal{V}_G$ instead of single vertices $v \in \mathcal{V}_G$.
Apart from this fundamental difference to 1-\acs{wl} bounded \acp{gnn}, $k$-\acp{gnn} additionally use two learnable parameter matrices $W^{(t)}, W_{\Gamma}^{(t)} \in \mathbb{R}^{d^{(t-1)} \times d^{(t)}}$ instead of the single $\Theta^{(t)}$ used in \acp{gcn}.
This separation of the parameter matrix effectively fulfills the same purpose as the \ac{mlp} in \ac{gin}; it makes the implicitly defined $k$-\acs{gnn} hash function injective.
This can be seen by realizing that a stack of $k$-\acs{gnn} convolution layers with $W_{\Gamma}^{(t)} = \mathbf{0}$ for all but the first layer simulates a \ac{mlp}.

As a final remark, note that the concept of ``neighborhood'' in \cref{eq:related:kgnn-layer} is slightly different from that used by the $k$-\acs{wl} algorithm (see \cfullref{defn:related:wlk-refine}):
$k$-\acs{gnn} considers a single $k$-multiset as a neighbor while $k$-\acs{wl} uses $k$-sets of $k$-multisets as neighbors.
We will get back to this difference later in \cref{sec:ltd:kgnn-limits}.

@@

The graph convolution approaches that we just looked at all produce a set of convolved feature vectors ${\left\{ Z^{(T)}[s_i] \right\}}_{i=1}^{m}$ after $T$ convolutional layers. % chktex 21
For 1-\acs{wl} bounded architectures like \ac{gcn} or \ac{gin}, there is one vector for each of the $m = {|\mathcal{V}_G|}$ vertices, while a $k$-\acs{wl} inspired architecture like $k$-\acs{gnn} produces $m \leq {|\mathcal{V}_G|}^{k}$ output vectors.
In order to solve the \ac{gcr} problem, those vector sets need to be combined into a final predicted class or regression value.
To do this, so-called \textit{graph pooling layers} are used.
Generally speaking there are two types of pooling approaches:
\begin{enumerate}[label=\textbf{\arabic*.}]
	\item \textbf{Hierarchical Pooling:}
		This type of pooling is similar to that found in \acp{cnn} on pictures where the resolution is iteratively reduced.
		Similarly hierarchical graph pooling iteratively \textit{coarsens} a graph every couple of convolutional layers.
		Graph coarsening works by merging vertices in a spectrum-preserving manner~\cite{Loukas2018}.
		This is done until a single merged vertex is left whose feature vector then represents the entire graph~\cite{Ying2018}.
	\item \textbf{Global Pooling}:
		Alternatively pooling can also be performed in a single step after the convolutional layers.
		This type of pooling takes all $m$ feature vectors and directly maps them to a single graph feature vector.
\end{enumerate}
In this thesis we will focus on global pooling.
The most simple global pooling approaches use static aggregation functions like component-wise $\min$, $\max$ or $\mean$ to produce a single graph feature vector $z_G \in \mathbb{R}^{d^{(T)}}$.
This vector is then fed into a standard \ac{mlp} to produce the final prediction.
The state-of-the-art graph pooling approaches go beyond static aggregation and try to incorporate structural information.
We will now briefly look at two such approaches.
